---
title: "Titanic Survival Prediction Model"
author: "Trevor Okinda"
date: ""
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

|                                              |     |
|----------------------------------------------|-----|
| **Student ID Number**                        | 134780 |
| **Student Name**                             | Trevor Okinda |
| **BBIT 4.2 Group**                           | C |
| **Project Name**                             | Titanic Survival Prediction Model |

# Setup Chunk

**Note:** the following KnitR options have been set as the global defaults: <BR> `knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = FALSE
)
```

# Understanding the Dataset (Exploratory Data Analysis (EDA))

## Loading the Dataset

### Source: 

The dataset that was used can be downloaded here: *\<https://www.kaggle.com/datasets/rahulsah06/titanic?resource=download\>*

### Reference:

*\<Avagyan, Z. (2017). Weather CSV [Data set]. Kaggle. https://www.kaggle.com/datasets/zaraavagyan/weathercsv\>\
Refer to the APA 7th edition manual for rules on how to cite datasets: <https://apastyle.apa.org/style-grammar-guidelines/references/examples/data-set-references>*

```{r Load dataset}
#Load dataset
TitanicData <-read.csv("titanic.csv", colClasses = c(
  PassengerId = "numeric",
  Pclass = "numeric",
  Name = "character",
  Sex = "factor",
  Age = "numeric",
  SibSp = "numeric",
  Parch = "numeric",
  Ticket = "character",
  Fare = "numeric",
  Cabin = "character",
  Embarked = "factor"
))

# Define levels for categorical columns
embarked_levels <- c("C", "Q", "S")
sex_levels <- c("female", "male")

# Update factor columns with levels
TitanicData$Embarked <- factor(TitanicData$Embarked, levels = embarked_levels)
TitanicData$Sex <- factor(TitanicData$Sex, levels = sex_levels)

# Display the dataset
View(TitanicData)
```